---
title: "Examine Land Use preferences in Porcupines"
output: html_notebook
---

In order to determine habitat suitability, we are working to identify which habitat types are found in the typical home ranges of our species so that we can use that information to build weights for habitat classes.

In lab, you used the "tabulate area" tool in ArcGIS pro to determine the count of the number of pixels of each habitat type that fell within each of the home range polygons you created for each observation of your focal species. You exported the table as a .csv. Today we'll see how to use those data to inform the next step of our project.

1. Set things up and suck in some data
```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Begin by loading the codes for landcover types and sourcing the function
```{r}
codes<-read.csv(here("Data", "LULC_codes.csv"))
source(here("Code", "Functions", "assignLandUse.R"))
source(here("Code", "Functions", "add_LULC_cases.R"))
```

Suck in the data on the landscape and use it to get average percent of landscape per class.

```{r}
NA_LULC<-read.csv(here("Data", "NA_LC_noOntario.csv"))
#drop to just first 3 columns; then add landcover codes
NA_LULC<-NA_LULC[,1:3]
colnames(NA_LULC)<-c("OBJECTID_1", "Value", "Count")
TotalArea<-sum(NA_LULC$Count)
NA_LULC<-assignLandUse(NA_LULC, codes)

NA_LULC$PercentLand<-round((100*(NA_LULC$Count/TotalArea)),2)
Land<-NA_LULC %>% select(Value, LandcoverNA, PercentLand)
```

Now write some code that steps through each file, sucks it in, processes it, and writes output to a separate directory.

Determine the list of files
```{r}
myfiles<-list.files(here("Data", "HR_no_PLCI"))
numfiles<-length(myfiles)
```
Get a list of species codes from the myfiles list
```{r}
Species<-str_sub(myfiles,1,4)
SpecNames<-c("Moose", "Ruffed grouse", "Porcupine", "Fisher",  "American woodcock", "Barred owl")
```
Create dataframe to hold results

```{r}
AvgPreference<-data.frame(Species = character(),
                          LandcoverNA = character(),
                          avg_use = numeric(),
                          avail = numeric(),
                          avg_pref = numeric(),
                          sd_use = numeric(),
                          sd_pref = numeric())
```

Start the loop

```{r}
for(i in 1:numfiles){
   #actual habitat use
   Use<-read.csv(here("Data", "HR_no_PLCI", myfiles[i]))
   
   #make sure each home range has all LULC categories
   Use<-add_LULC_cases(Use) 
   #assign land use codes to each category
   Use<-assignLandUse(Use, codes) 
   

#determine percent land use
Percent_summary<-Use %>% group_by(OBJECTID_1) %>% summarize(
   tot_area = sum(Area)
)
#now join by OBJECTID_1
Use<-left_join(Use, Percent_summary)
Use$PercentAreaUse<-round((100*(Use$Area/Use$tot_area)),2)

#now join with Land
Use<-left_join(Use, Land)

#calculate use per availability
Use$UperA<-Use$PercentAreaUse/Use$PercentLand

#Now get summary

AvgPrefSummary<-Use %>% group_by(LandcoverNA) %>% summarize(
   avg_use = mean(PercentAreaUse, na.rm = T),
   avail = mean(PercentLand),
   avg_pref = mean(UperA, na.rm = T),
   sd_use = sd(PercentAreaUse, na.rm = T),
   sd_pref = sd(UperA, na.rm = T)
)
AvgPrefSummary$Species<-Species[i]
AvgPrefSummary<-AvgPrefSummary %>% relocate(Species, .before = LandcoverNA)
#put data into placeholder
AvgPreference<-rbind(AvgPreference, AvgPrefSummary)

#And now plot, colored to indicate use
P<-ggplot(AvgPrefSummary, aes(LandcoverNA, avg_pref, label = round(avg_pref, 2)))+
   geom_hline(yintercept = 1)+geom_point(aes(color = cut(avg_pref, c(0,1,10))), size = 2)+
   scale_color_manual(name = "Use per availablilty",
                      values = c("(0,1]" = "red",
                                 "(1,10]" = "dark green"),
                      labels = c("less than 1", "greater than 1"))+
geom_text(vjust = -1)+
   labs(x = "Land use category", y = "Use per availability", title = paste0("Habitat preference - ", SpecNames[i]))+
   coord_flip()

assign(paste0(Species[i],"_pref_plot"),P)
figname<-paste0(Species[i], "_pref_plot.png")
ggsave(here("Output", figname), device = "png")
}
```
Now run for red-backed salamander, which needs a different approach since each individual home range is smaller than a pixel.

```{r}
#actual habitat use
   Use<-read.csv(here("Data", "HR_PLCI", "PLCI_80m2_LULC.csv"))
   
   #assign land use codes to each category
   Use<-assignLandUse(Use, codes) 
   
#now summarize num observations of each land use code
Percent_summary<-Use %>% group_by(LandcoverNA) %>% summarize(
   num_hrs = length(Value))

#determine total number of HRs
Tot_HRS<-sum(Percent_summary$num_hrs)

#now get percent of observations
Percent_summary$Obs_percent<-round((100*(Percent_summary$num_hrs/Tot_HRS)),2)

#now join with Land
Percent_summary<-left_join(Percent_summary, Land)

#calculate use per availability
Percent_summary$UperA<-Percent_summary$Obs_percent/Percent_summary$PercentLand

#Now make compatible with main data frame
colnames(Percent_summary)<- c("LandcoverNA", "num_hrs", "avg_use", "Value", "avail", "avg_pref")

AvgPrefSummary<-Percent_summary
AvgPrefSummary$Species<-"PLCI"
AvgPrefSummary<-AvgPrefSummary %>% relocate(Species, .before = LandcoverNA)
AvgPrefSummary<-AvgPrefSummary %>% select(Species, LandcoverNA, avg_use, avail, avg_pref)

AvgPrefSummary$sd_use<-NA
AvgPrefSummary$sd_pref<-NA

  

#put data into placeholder
AvgPreference<-rbind(AvgPreference, AvgPrefSummary)

#And now plot, colored to indicate use
PLCI_pref_plot<-ggplot(AvgPrefSummary, aes(LandcoverNA, avg_pref, label = round(avg_pref, 2)))+
   geom_hline(yintercept = 1)+geom_point(aes(color = cut(avg_pref, c(0,1,10))), size = 2)+
   scale_color_manual(name = "Use per availablilty",
                      values = c("(0,1]" = "red",
                                 "(1,10]" = "dark green"),
                      labels = c("less than 1", "greater than 1"))+
geom_text(vjust = -1)+
   labs(x = "Land use category", y = "Use per availability", title = "Habitat preference - Red backed salamander")+
   coord_flip()

ggsave(here("Output", "PLCI_pref_plot.png"), device = "png")
```

Now save data file
```{r}
write.csv(AvgPreference, here("Output", "All_species_hab_preference.csv"))
```






