---
title: "Fix categories"
output: html_notebook
---
This is code to quickly go through a series of files and fix the Landcover category names to be consistent among users

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Get a list of the files in the directory

Suck in one file and fix it

```{r}
myFile<-read.csv(here("Data", "HR", "PLCI_LULC_table.csv"))
New_cols<-colnames(myFile)
New_cols[7:9]<-c("LandcoverNA", "NLCD_US", "LandcoverUS")
New_cols[5]<-"orig_area"
New_cols[6]<-"Area"
colnames(myFile)<-New_cols
```
Now strip out just the columns I want
```{r}
myFile<-myFile %>% select(OID_, OBJECTID_1, Value, Count, Area)
```
And write
```{r}
write.csv(myFile, here("Data", "Intermediate", "PLCI_LULC_int.csv"))
```
Now next one
```{r}
myFile<-read.csv(here("Data", "HR","ALAL_LULC_table.csv"))
```
Inspect the columns
```{r}
New_cols<-colnames(myFile)
```
Now strip out just the columns I want
```{r}
myFile<-myFile %>% select(OID_, OBJECTID_1, Value, Count, Area)
```
And write
```{r}
write.csv(myFile, here("Data", "Intermediate", "ALAL_LULC_int.csv"))
```

And the next
```{r}
myFile<-read.csv(here("Data", "HR","BOUM_LULC_table.csv"))
colnames(myFile)
```
Now strip out just the columns I want
```{r}
myFile<-myFile %>% select(OID_, OBJECTID_1, Value, Count, Area)
```

And write
```{r}
write.csv(myFile, here("Data", "Intermediate", "BOUM_LULC_int.csv"))
```

And the next
```{r}
myFile<-read.csv(here("Data", "HR","STVA_LULC_table.csv"))
colnames(myFile)
```
Now strip out just the columns I want
```{r}
myFile<-myFile %>% select(OID_, OBJECTID1, Value, Count, Area)
```

And write
```{r}
write.csv(myFile, here("Data", "Intermediate", "STVA_LULC_int.csv"))
```







